HMM Model development

### 2020-05-18 dev/hmm_model_devel.Rmd
Towards better model wrapping, posterior predictive checks

improving_data <- data.frame(day = data$observed$times, patient = data$observed$series, improving = data$true$true_improving) %>% filter(improving)

data.frame(day = data$observed$times, patient = data$observed$series, breathing = data$true$true_base_states) %>%
  mutate(breathing = factor(breathing, 1:data$observed$N_states_observed, ordered = TRUE)) %>%
  ggplot(aes(x = day, y = patient, fill = breathing)) + geom_tile(width = 1, height = 0.5) + geom_tile(data = improving_data, width = 1, height = 0.1, fill = "gray") + ggtitle("True - gray lines for improving")

data.frame(day = data$observed$times, patient = data$observed$series, breathing = data$observed$obs_states) %>%
  mutate(breathing = factor(breathing, 1:data$observed$N_states_observed, ordered = TRUE)) %>%
  ggplot(aes(x = day, y = patient, fill = breathing)) + geom_tile(width = 1, height = 0.5) + ggtitle("Observed")
```

@@

single_code <- make_stancode_hmm(data$observed$formula, 
                                 serie_data = data$observed$serie_data, 
                                 rate_data = data$observed$rate_data, 
                                 hidden_state_data = data$observed$hidden_state_data, 
                                 initial_states = data$observed$initial_states,
                                 observed_state_data = data$observed$observed_state_data,
                                 prior= data$observed$prior)

@@

standata <- make_standata_hmm(data$observed$formula, 
                              serie_data = data$observed$serie_data, 
                              rate_data = data$observed$rate_data, 
                              hidden_state_data = data$observed$hidden_state_data, 
                              initial_states = data$observed$initial_states,
                              observed_state_data = data$observed$observed_state_data,
                              sensitivity_low_bound = data$observed$sensitivity_low_bound)

@@

```{r}
noisy_states_in_sbc <- FALSE
sbc_generator <- function() {
  data <- hmm_simulator(50, 20, 3, use_noisy_states = noisy_states_in_sbc)
  standata <- make_standata_hmm(data$observed$formula, 
                                serie_data = data$observed$serie_data, 
                                rate_data = data$observed$rate_data, 
                                hidden_state_data = data$observed$hidden_state_data, 
                                initial_states = data$observed$initial_states,
                                observed_state_data = data$observed$observed_state_data,
                                sensitivity_low_bound = data$observed$sensitivity_low_bound)

  list(observed = standata,
       true = data$true)
}

sbc_res <- sbc(model_single, generator = sbc_generator, N_steps = 200, control = list(adapt_delta = adapt_delta))

@@

```{r}
library(brms)
```

```{r}
```

@@

test_data <- data.frame(a = c(1,2,3), b = c(0.5,1.0,1.5), x1 = c("A"), x2 = c(0,1,2))
f <- a ~ 0 + (0 + x2 | a | x1)
# make_stancode(f, data = test_data)
#brms::make_standata(f, data = test_data)
xx <- brms::brmsformula(f)

xx$resp

stan_llh.rate_hmm <- function( ... ) {
  ""
}

.family_rate_hmm <- function() {

}

ff <-

make_stancode(a ~ 0 + x2, family = ff, data = test_data)

>>>>

improving_data <- data.frame(day = data$observed$serie_data$.time, patient = data$observed$serie_data$.serie, improving = data$true$true_improving) %>% filter(improving)

data.frame(day = data$observed$serie_data$.time, patient = data$observed$serie_data$.serie, breathing = data$true$true_base_states) %>%
  mutate(breathing = factor(breathing, 1:nrow(data$observed$observed_state_data), ordered = TRUE)) %>%
  ggplot(aes(x = day, y = patient, fill = breathing)) + geom_tile(width = 1, height = 0.5) + geom_tile(data = improving_data, width = 1, height = 0.1, fill = "gray") + ggtitle("True - gray lines for improving")

data.frame(day = data$observed$serie_data$.time, patient = data$observed$serie_data$.serie, breathing = data$observed$serie_data$.observed) %>%
  mutate(breathing = factor(breathing, 1:nrow(data$observed$observed_state_data), ordered = TRUE)) %>%
  ggplot(aes(x = day, y = patient, fill = breathing)) + geom_tile(width = 1, height = 0.5) + ggtitle("Observed")
```

@@

single_code <- make_stancode_hmm(data$observed)

@@

standata <- make_standata_hmm(data$observed)

@@

```{r}
data <- hmm_simulator(5, 10, 2, use_noisy_states = TRUE)

bfit <- brmhmm(data$observed)

summary(bfit)

xx <- brms::posterior_epred(bfit)
dim(xx)
```

```{r}
noisy_states_in_sbc <- TRUE
sbc_generator <- function() {
  data <- hmm_simulator(40, 25, 2, use_noisy_states = noisy_states_in_sbc)
  standata <- make_standata_hmm(data$observed)

  list(observed = standata,
       true = data$true,
       observed_raw = data$observed)
}

model_sbc <-  stan_model(model_code = make_stancode_hmm(sbc_generator()$observed_raw))

sbc_res <- sbc(model_sbc, generator = sbc_generator, N_steps = 50, control = list(adapt_delta = adapt_delta))

@@

## Apply to Grein et al. dataset

```{r}
data_grein <- readxl::read_excel(here::here("public_data", "Grein_et_al.xlsx"), na = c("")) %>%
  select(-Remdesivir_Last_Day) %>%
  pivot_longer(starts_with("Day_"), names_prefix = "Day_", names_to = "day", values_to = "breathing") %>%
  filter(!is.na(breathing)) %>%
  transmute(
    patient = Patient,
    breathing = breathing,
    day = as.integer(day)) %>%
  group_by(patient) %>%
  mutate(breathing = if_else(breathing == "ND", breathing[which.max(day[breathing %in% c("AA", "LoO2", "HiO2", "NIPPV", "MV", "ECMO")])], breathing)) %>%
  ungroup() %>%
  mutate(
    breathing = factor(breathing, levels = c("Discharge", "AA", "LoO2", "HiO2", "NIPPV", "MV", "ECMO", "Death"), ordered = TRUE)
  )

@@
